---
title: "HomeDepot First Data Exploration"
author: "Brian Carter"
date: "19 February 2016"
output: html_document
---

```{r setup,include=FALSE}

#Set some global options for r chunks. 
##########Options####################
#Three possibilites, code is evaluate, code is printed, output is displayed. 
#
# include=FALSE - the code is evaluated but neither the code displayed or output. 
# echo=FALSE - display output, do not display code chunk
# eval=FALSE - display code, do not evaluate
#


# results = 'asis' for displaying out of table when formatted.



#Want R chunk to evaluate, but not included for client facing document. 

knitr::opts_chunk$set(cache=FALSE,fig.height=4,fig.width=6,fig.align='center',echo=FALSE,eval=TRUE,warning=FALSE,message=FALSE,error=FALSE) #(fig.path='figure/') ##background doesn't work

#Caching gives some funny results but it can result in faster loading times, when writing up the document. This is to do with the := assignment in data.table. Knitr doesn't recognise, therefore always use <-
```


```{r}

##Libraries required
#Data work 
require(data.table) #Working with large files
require(xlsx)       #Loading and saving .xlsx files 
require(plyr)   #Always load in this order plyr, dpply, lubridate - dpplyr overrides some of methods in plyr. 
require(dplyr) #Use require as it will give an error message if the package doesn't exist
require(lubridate) #used for working with data information. 
require(reshape2)  #used for melting 


#Formating and printing 
#install.packages("devtools")
#devtools::install_github("adletaw/captioner")   #Nice library for numbering and captioning tables in conjunction with knitr and pandoc
require(pander)   	#for creating nice output tables.
require(captioner)

#Set up the figure and table numbering
fig_nums<-captioner()
tab_nums<-captioner(prefix = "Table")

#Using pryr abbreviate how to call fig_nums function 
require(pryr)
citefig<-pryr::partial(fig_nums,display="cite")
citetab<-pryr::partial(tab_nums,display="cite")

#Turn off caption.prefix as allow captioner to handle this. 
panderOptions('table.caption.prefix', '')
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)
```

An initial analysis of the data in the *train.csv* file from the Home Depot Competion. (Where new variables created, prefix with a *c.*) There are four files of interest for exploration *(train.csv,test.csv,product\_descriptions.csv,attributes.csv,sample\_submission)*. 

```{r,cache=FALSE}
#Load the data 
hdTrain<-data.table::fread("../input/train.csv",header=TRUE,encoding="Latin-1") 
hdTest<-data.table::fread("../input/test.csv",header=TRUE) 
hdAtt<-data.table::fread("../input/attributes.csv",header=TRUE)

#Not really need here
hdSS<-data.table::fread("../input/sample_submission.csv",header=TRUE) 
hdDesc<-data.table::fread("../input/product_descriptions.csv",header=TRUE)

```

The *sample\_submission.csv* file and the *test.csv* file have the same number of rows, `r nrow(hdSS)`. 

# train.csv and test.csv Overlap. 

There are `r nrow(hdTrain)` rows in the *train.csv* data and `r nrow(hdTest)` in the *test.csv* data. Both have the same columns [`r names(hdTrain)`] bar the **relevance** score which is not included in the *test.csv* data. Examining the column **product_id** there are `r length(unique(hdTrain[,product_uid]))` unique values in the *train.csv* data and and `r length(unique(hdTest[,product_uid]))` in the *test.csv* data. There are total of `r length(intersect(hdTrain[,product_uid],hdTest[,product_uid]))` **product_uid** values that intersect. This is displayed below in a venn diagram. 

```{r,fig.width=4,fig.height=3,echo=TRUE}
#Count unique *product_uid*, the PK across datasets.
pidTrain = unique(hdTrain[,product_uid])
pidTest = unique(hdTest[,product_uid])
pidAtt = unique(hdAtt[,product_uid])
pidDesc = unique(hdDesc[,product_uid])

#Create venn diagram to represent the relationship between Product IDs across both. 

#install.packages("VennDiagram")
library(VennDiagram)

grid.newpage()

venn.plot1<-draw.pairwise.venn(
	length(pidTrain),length(pidTest),length(intersect(pidTrain,pidTest)),
	category = c("Train product_uid", "Test product_uid"),
	lty = rep("blank",2),
	fill = c("light blue", "pink"),
	alpha = rep(0.5, 2),
	cat.pos = c(0,0),
	cat.dist = rep(0.025, 2))

grid.draw(venn.plot1)
```

There are a total of `r length(pidTrain)` unique **product_uid** in the *train* and `r length(pidTest)` in *test*. Of these only `r length(intersect(pidTrain,pidTest))` are common. The total number of unique *product_uids* across both the *train* and *test* is `r length(union(pidTest,pidTrain))`. This equals the total number of **product_uid** and **rows** in the *product_descriptions.csv* file,(`r nrow(hdDesc)==length(unique(hdDesc[,product_uid]))`). 

In the *attributes.csv* file, there are a total of `r nrow(hdAtt)` **rows** and a total of `r length(unique(hdAtt[,product_uid]))` unique **product\_uid**. The interesection of **product_uid** across the three files *(train,test,attributes)* is displayed in the venn diagram below. 

```{r,fig.width=4,fig.height=3}
grid.newpage()
venn.plot2<-draw.triple.venn(area1 = length(pidTest), area2 = length(pidTrain), area3 = length(pidAtt), 
														 n12 = length(intersect(pidTest,pidTrain)), 
														 n23 = length(intersect(pidTrain,pidAtt)),
														 n13 = length(intersect(pidTest,pidAtt)),
														 n123 =length(intersect(pidAtt,intersect(pidTest,pidTrain))),
														 category = c("test", "train", "attribute"),
														 lty = "blank",
														 fill = c("skyblue", "pink1", "mediumorchid"))
grid.draw(venn.plot2)
rm(venn.plot1,venn.plot2)
detach("package:VennDiagram",unload=TRUE)
#use search() to check if unloaded


```

There is one value in *attributes.csv* file that is not in either the *Train* or *Test* files. On investigating this there are `r nrow(hdAtt[is.na(product_uid),])` rows that do not have a *product_uid* value. These rows can be removed. 

```{r,echo=TRUE}
hdAtt<-hdAtt[!is.na(product_uid),]
```

# train.csv Data Exploration

The target variable is the **relevance** score. In the supplied *relevance_instructions.cocx* file supplied to the human graders it instructs to provide a relevance grade of either;

* Irrelevant (1)
* Partially or somewhat relevant (2)
* Perfect match (3). 

Further more in the competition data it states that *Each pair [search_term,product] was evaluated by at least three human raters. The provided relevance scores are the aveerage value of the ratings*. There are a total of `r length(unique(hdTrain[,relevance]))` unique values for **relevance** in the *train* data. A bar-chart of their frequency is displayed below. 

```{r,echo=TRUE}
library(ggplot2)
ggplot(hdTrain,aes(relevance))+geom_bar()+labs(title="Frequency of 13 individual relevance scores",x="relevance score",y="row count")
```

Note there are a set of **relevance** scores that occur with low frequenices (1.25,1.5,1.75,2.25,2.5,2.75), thus indicating that perhaps more than 3 or less than 3 human evaluators appraised that row. 

```{r,echo=TRUE}
#Create frequency table for printing
mytable<-data.frame(table(hdTrain$relevance))
names(mytable)<-c("Rel. Score","Frequency")
tab_nums("relFreq","Count of frequency of the relevance score")
```
```{r,echo=TRUE,align}
pander(mytable,caption=tab_nums("relFreq"),split.table=60,justify=c("left"),table.alignment="center")
```

`r citetab("relFreq")` displays the count of the frequecies of the **relevance** scores. There are `r nrow(hdTrain[relevance %in% c(1.25,1.5,1.75,2.25,2.5,2.75),])` rows in the *train* data that relate to these infrequent values, representing `r  round(nrow(hdTrain[relevance %in% c(1.25,1.5,1.75,2.25,2.5,2.75),])/nrow(hdTrain),6)*100`% of the total rows. It seems these are anomalies that could be removed. For the remainder of the exploratory analysis they are removed. 

```{r,echo=TRUE}
hdTrain<-hdTrain[!relevance %in% c(1.25,1.5,1.75,2.25,2.5,2.75),]
```

The average **relevance** score is examined by nummber of times a **product_uid** occurs in the dataset.  


```{r,echo=TRUE}
hdTrain %>% 
	group_by(product_uid) %>%
	summarize(count.PUID=.N,avg.Relevance=mean(relevance,na.rm=TRUE)) %>%
	ggplot(aes(x=count.PUID,y=avg.Relevance,group=1)) +
	geom_jitter(alpha=0.2) +
	scale_y_continuous(breaks = round(seq(1,3,by = 0.333),1))+
	labs(title="Average Relevance by count of product_uid")
```

As expected the variation of the **relevance** decreases as the number of times a *product_uid* occurs in the dataset. The **relevance** score is examined in the context of the number of parameters (words) in a **search_term**. 

```{r,echo=TRUE}
#count the number of search terms
hdTrain<-hdTrain[,c.count_search_term:=sapply(gregexpr("\\W+", search_term), length) + 1,]

#Create averages for annotating the plot
avgST<-round(mean(hdTrain$c.count_search_term),2)
avgSTGroup<-hdTrain %>% group_by(as.factor(relevance)) %>% summarise(groupMean=round(mean(c.count_search_term),1))
names(avgSTGroup)=c("relevance","grpMean")

#Create plot, limit y-axis,add overlay boxplot, annotate with meanby group and text. 
plotText<-paste("The mean relevance is ",avgST,"\ny-axis has been reduced.\n max number of  search_terms =",max(hdTrain$c.count_search_term))

hdTrain %>% 
	ggplot(aes(x=factor(relevance),y=c.count_search_term)) +
	geom_jitter(alpha=0.05) +
	coord_cartesian(ylim=c(1,12)) +
	geom_boxplot(color = "red", outlier.colour = NA, fill = NA)+
	geom_text(data = avgSTGroup, aes(x = relevance, y = grpMean, label = grpMean), size = 5, vjust = -0.25,colour="green")+
	annotate("text",label=plotText,x=2,y=11,size=3,colour="black")+
	labs(title="Average relevance by number search_terms",x="relevance",y="Count of search_terms")
```


```{r,eval=FALSE}
#Facetting code not used

ggplot(hdTrain, aes(x=c.count_search_term,colour=as.factor(relevance)))+geom_freqpoly(binwidth = 1)

ggplot(hdTrain, aes(x=c.count_search_term,colour=as.factor(relevance)))+geom_histogram(aes(group=as.factor(hdTrain$relevance),fill=as.factor(relevance)),binwidth = 1,alpha=0.3)

ggplot(dfs, aes(x=values)) + geom_density(aes(group=ind, colour=ind, fill=ind), alpha=0.3)

plot<-ggplot(hdTrain, aes(x=c.count_search_term,fill=as.factor(c.count_search_term)))+geom_histogram(binwidth = 1)+theme(legend.position='none')
plot+facet_wrap(~ relevance,nrow=1)

```

The Jaccard Similarity Coefficent between {**search_term**} and the {**product_title**} is caculated and plotted against the **relevance score**. 

$$JC(A,B) = \frac{|A \cap B|}{|A \cup B|} = \frac{|A \cap B|}{|A|+|B| - |A \cap B|}  $$ 


```{r,echo=TRUE}
#Function for calculating JC
calcJacCoef<-function(A,B){
	
	#Using library(sets) for calculating Jacard Coefficent.
	### Note it the warning messages: %>% is masked from package:dplyr, set is masked from package:data.table
	### Deatch it after its use otherwise gets annoying. 
	require(sets)
	
	A<-gset(strsplit(A, " ")[[1]])
	B<-gset(strsplit(B, " ")[[1]])
	return(gset_similarity(A,B,"Jaccard"))
	
	detach("package:sets",unload=TRUE)
	#use search() to check if unloaded
}

#Calculate the Jaccard Coefficient between the *search_term&* and *product_title* for each row in hdTrain
#Takes time to calculate. 
hdTrain<-hdTrain[,search_terms:=tolower(search_term)]
hdTrain<-hdTrain[,product_title:=tolower(product_title)]
hdTrain<-hdTrain[,c.Jaccard.ST_PT := calcJacCoef(search_term,product_title),by = 1:nrow(hdTrain)]

detach("package:sets",unload=TRUE)
#use search() to check if unloaded

#Create averages for annotating the plot
avgJC<-round(mean(hdTrain[,c.Jaccard.ST_PT]),2)
avgJCGroup<-hdTrain %>% group_by(relevance) %>% summarise(groupMean=round(mean(c.Jaccard.ST_PT),2))
names(avgJCGroup)=c("relevance","grpMean")


hdTrain %>% 
	ggplot(aes(x=factor(relevance),y=c.Jaccard.ST_PT)) +
	geom_jitter(alpha=0.05) +
	geom_boxplot(color = "yellow", outlier.colour = NA, fill = NA)+
	geom_text(data = avgJCGroup, aes(x = factor(relevance), y = grpMean, label = grpMean), size = 5, vjust = 0,colour="green")+
	labs(title="Average Jaccard Coefficient by Relevance Score",x="relevance",y="Jaccard Coefficent")
```

#Bring in attributes.csv

There are `r nrow(hdAtt)` rows in the *attributes.csv* file. These relate to `r length(unique(hdAtt[,product_uid]))` unique products. The plot below show displays a the **relevance** by those products in the *attributes* and those that are not. 

```{r}

hdTrain<-hdTrain[product_uid %in% unique(hdAtt[,product_uid]),c.hasAttribute:="Yes"]
hdTrain<-hdTrain[is.na(c.hasAttribute),c.hasAttribute:="No"]

hdTrain %>% 
	ggplot(aes(x=factor(c.hasAttribute),y=relevance)) +
	geom_jitter(alpha=0.05) +
	geom_boxplot(color = "red", outlier.colour = NA, fill = NA)+
	labs(title="Compare relevance by product_uid in Attribute file",x="Product has Attributes",y="relevance")

```


In the attributes file there are `r length(unique(hdAtt[,tolower(name)]))` different categories(**name**) of *attributes*. 

```{r}
#Create frequency table for printing
mytable<-data.table(table(tolower(hdAtt$name)))
names(mytable)<-c("Attribute.name","Freq.")
mytable<-mytable[with(mytable,order(-Freq.)),]
tab_nums("attNameFreq","Count of Top 10 Freq. of categories in attribute.csv")
```
```{r}
pander(mytable[1:10,],caption=tab_nums("attNameFreq"),split.table=60,justify=c("left"),table.alignment="center")
```


```{r}
mytableColor<-mytable[Attribute.name %like% "colo*r",]
mytableBrand<-mytable[Attribute.name %like% "brand",]
tab_nums("attNameFreqColor","Count of Top 10 Freq. of Colour categories")
tab_nums("attNameFreqBrand","Count of Top 10 Freq. of Brand categories")
```

There are `r nrow(mytableColor)` **name** that refer to a   <span style="color:red">color</span>, aspect of the product in the *attribute file*. Similarly there are `r nrow(mytableBrand)` that refer to  <span style="color:red">brand</span> aspect. A sorted count of the top 10 for each are are displayed below. 

```{r}
pander(mytableColor[1:10,],caption=tab_nums("attNameFreqColor"),split.table=60,justify=c("left"),table.alignment="center")
pander(mytableBrand[1:10,],caption=tab_nums("attNameFreqBrand"),split.table=60,justify=c("left"),table.alignment="center")
```


The **name=mrg brand name** appears to be the standard for Brand attributes. For color the information is spread acorss multiple fields. Create two flags in the *Train* indicating if a **product_uid** has brand and color attribute information. 
	
```{r,echo=TRUE}
#First change hdAtt tolower
hdAtt<-hdAtt[,name:=tolower(name)]
	
hdTrain<-hdTrain[product_uid %in% unique(hdAtt[name %like% "colo*r",product_uid]),c.hasColorAtt:="Yes"]
hdTrain<-hdTrain[is.na(c.hasColorAtt),c.hasColorAtt:="No"]
	
hdTrain<-hdTrain[product_uid %in% unique(hdAtt[name == "mfg brand name",product_uid]),c.hasBrandAtt:="Yes"]
hdTrain<-hdTrain[is.na(c.hasBrandAtt),c.hasBrandAtt:="No"]
```
	
```{r}
hdTrain[c.hasAttribute=="Yes",] %>%
	ggplot(aes(x=factor(c.hasBrandAtt),y=relevance)) +
	geom_jitter(alpha=0.05) +
	geom_boxplot(color = "blue", outlier.colour = NA, fill = NA)+
	labs(title="Compare relevance by product_uid in attributes file \n and has Attribute.name==mfg brand name",x="Product has mfg brand name in    attributes",y="relevance")
```
	
	
```{r}
hdTrain[c.hasAttribute=="Yes",] %>%
	ggplot(aes(x=factor(c.hasColorAtt),y=relevance)) +
	geom_jitter(alpha=0.05) +
	geom_boxplot(color = "green", outlier.colour = NA, fill = NA)+
	labs(title="Compare relevance by product_uid in attributes file \n and has Attribute.name of type color",x="Product has color attributes",y="relevance")
```
	
	
	


